---
title: "Model ~ Treatment + Animal.ID"
output: html_document
---
##Loading Libraries
```{r echo=TRUE, message=FALSE}
library(SummarizedExperiment)
library(edgeR)
library(limma)
library(ggplot2)
library(dplyr)

sexp <- readRDS(gzcon(url("http://salomon24.scripps.edu/~ryan/Cyno-RNASeq-SummarizedExperiment.RDS")))
sample_table <- as.data.frame(colData(sexp))
gene_info <- as.data.frame(mcols(sexp))
count_matrix <- assay(sexp)
```

##Normalization
```{r}
total_counts <- colSums(count_matrix)
nf <- calcNormFactors(count_matrix, lib.size=total_counts, method="TMM")
normalized_total_counts <- total_counts * nf
```

##Filter non expressed genes
```{r}
mean_log_cpm <- aveLogCPM(count_matrix, normalized_total_counts)
filter_threshold <- 1
keep_genes <- mean_log_cpm >= 1
filtered_count_matrix <- count_matrix[keep_genes,]
filtered_gene_info <- gene_info[keep_genes,]
```

#Model ~ Treatment

##Fitting linear model
```{r}
design <- model.matrix(~ Treatment + Animal.ID, data=sample_table)
v <- voom(filtered_count_matrix, design, lib.size=normalized_total_counts, plot=TRUE)
v$genes <- filtered_gene_info
fit <- lmFit(v, design)
fit <- eBayes(fit, robust=TRUE)
results <- topTable(fit, coef="TreatmentIFNg", adjust.method="BH", n=Inf)
```

##MDS plot
```{r}
mds <- data.frame(plotMDS(v)[c("x", "y")])
mds <- cbind(mds, sample_table)
p <- ggplot(mds) +
  aes(x=x, y=y) +
  xlab("PC1") + ylab("PC2") +
  geom_point(size=3) +
  coord_fixed(ratio=1) +
  ggtitle("Sample MDS Plot")
p + aes(color=Animal.ID, shape=Treatment)
```

#Removing bad animal
```{r}
bad_animal <- "CN8351"
selected_samples <- !(sample_table$Animal.ID %in% bad_animal)
good_sample_table <- droplevels(sample_table[selected_samples,])
good_filtered_count_matrix <- count_matrix[keep_genes,selected_samples]
good_normalized_total_counts <- normalized_total_counts[selected_samples]
```

#Fitting new linear model
```{r}
good_design <- model.matrix(~ Treatment + Animal.ID, data=good_sample_table)
good_v <- voom(good_filtered_count_matrix, good_design, lib.size=good_normalized_total_counts, plot=TRUE)
good_v$genes <- filtered_gene_info
good_fit <- lmFit(good_v, good_design)
good_fit <- eBayes(good_fit, robust=TRUE)

good_results <- topTable(good_fit, coef="TreatmentIFNg", adjust.method="BH", n=Inf)
```

##MA Plot
```{r}
ggplot(arrange(results, desc(P.Value))) +
  aes(x=AveExpr, y=logFC,
      color=ifelse(adj.P.Val <= 0.1, "FDR <= 10%", "FDR > 10%")) +
  geom_point(size=0.1) +
  scale_color_hue(name="Significance") +
  theme(legend.justification=c(1,1), legend.position=c(1,1)) +
  ggtitle("MA Plot, IFNg vs Control")
```

##New MA Plot
```{r}
ggplot(arrange(good_results, desc(P.Value))) +
  aes(x=AveExpr, y=logFC,
      color=ifelse(adj.P.Val <= 0.1, "FDR <= 10%", "FDR > 10%")) +
  geom_point(size=0.1) +
  scale_color_hue(name="Significance") +
  theme(legend.justification=c(1,1), legend.position=c(1,1)) +
  ggtitle("MA Plot, IFNg vs Control")
```

##P-value distribution
```{r}
ggplot(results) +
  aes(x=P.Value) +
  geom_histogram(aes(y=..density..), binwidth=0.025, boundary=0) +
  geom_hline(yintercept=1) +
  ggtitle("P-value distribution for Control vs Treatment")
```
##New P-value distribution
```{r}
ggplot(good_results) +
  aes(x=P.Value) +
  geom_histogram(aes(y=..density..), binwidth=0.025, boundary=0) +
  geom_hline(yintercept=1) +
  ggtitle("P-value distribution for Control vs Treatment")
```

##Significant Differentially expressed genes w.r.t. Treatment (FDR <= 10%)
###Model
```{r}
table(results$adj.P.Val <= 0.1)
```
###New Model
```{r}
table(good_results$adj.P.Val <= 0.1)
```

##Significant Differentially expressed genes w.r.t. Animal.ID (FDR <= 10%)
###Model
```{r}
animal.results <- topTable(fit, coef=c("Animal.ID6C63", "Animal.ID6C7", "Animal.ID6C84", "Animal.ID7C37", "Animal.ID8C8", "Animal.IDCN7314", "Animal.IDCN7875"), n=Inf)
table(animal.results$adj.P.Val <= 0.1)
```
###New Model
```{r}
animal.good_results <- topTable(fit, coef=c("Animal.ID6C63", "Animal.ID6C7", "Animal.ID6C84", "Animal.ID7C37", "Animal.ID8C8", "Animal.IDCN7314", "Animal.IDCN7875"), n=Inf)
table(animal.good_results$adj.P.Val <= 0.1)
```